A novel route to the spontaneous formation of porous crystals via viscoelastic phase separation

### 2016-03-15 nature/crystal_gel2.tex
Dear all,
Thank you for your efforts. I made corrections to the main text, the Supplementary Information, and the response letter. Please check them carefully. There are also some places we need to fix. Thanks!! Hajime

To gain additional insight, in Fig.~\ref{fig:phase_separation}d we plot the average local volume fraction as a function of the number of neighbours.
\textcolor{red}{The volume fraction is obtained by computing the Voronoi diagram of each configuration,
which uniquely assigns a volume to each colloidal particle.}
For all state points, we can see that the average volume fraction of a colloidal
particle increases with its number of neighbours. Approaching close packing (12 neighbours), we clearly see two families of curves. The state point
at $\phi\approx 0.33$ and $c_p=0.38$ mg/g reaches close packing at a volume fraction of around $\sim 68\%$, which is in reasonable agreement with the
volume fraction of the stable crystal phase~\cite{ilett1995phase}. All other state points, instead, reach close packing at a volume fraction $\sim 64\%$, which is
indeed close to the volume fraction of the attractive glass state~\cite{pham2002multiple}. 
%(\textbf{the absolute values of these volume fraction are not totally convincing, probably because of a not optimal estimate of the size of the colloids. We should run the code of Mathieau with size detection}). 
This result clearly indicates that the state point at $\phi\approx 0.33$ and $c_p=0.38$ mg/g follows a different arrest mechanism, in which phase separation
is arrested by crystallization and not by glassiness. Although our estimation of the absolute volume fraction may involves systematic errors of a few \%, the trend should be robust~\cite{Poon2012}.

@@

\textcolor{red}{Next we analyse in detail the crystallization trajectory, distinguishing between the different process responsible for the
crystallization of the phase separated network.
Supplementary Movie 2 shows the confocal microscopy video for the crystallizing sample at $\phi\approx 0.33$ and $c_p=0.38$ mg/g,
where the process of crystal growth by direct attachment of gas particles can be observed directly. 
From the trajectory, we extract slabs of 10 particle diameter thickness for configurations at different times, shown in Fig.~\ref{fig:transitions}a. 
Particles are coloured according to their phases (gas, liquid, and crystal). For the classification criteria see the \emph{Methods} section. The figure shows the first nucleation 
events inside the liquid network (left panel). At the same time liquid regions that have not crystallized start evaporating, and the gas phase contributes to the growth of nuclei (middle panel). Finally the different nuclei coalesce (right panel). Note that the network pattern changes between $t=2055\,\tau_B$ (left panel) and $t=5577,\tau_B$ (middle panel): not all liquid
branches of the network crystallize due to evaporation. The crystal-gel structure instead persists, as seen by comparing $t=5577,\tau_B$ (middle panel) and $t=12098,\tau_B$ (right panel).}

\textcolor{red}{
Next we show in Fig.~\ref{fig:transitions}b the fraction of particles in each different phase for the crystallization trajectory at
$\phi\approx 0.33$ and $c_p=0.38$ mg/g, after the liquid-gas phase separation has occurred.
For the classification criteria see \emph{Methods}.
The process of crystallization is characterized 
by a steep decrease in liquid particles, as they transform into small crystalline nuclei inside the liquid domains. This decrease
is then accompanied by an increase in the fraction of gas particles: as the first crystals start to reach the gas phase, liquid particles evaporate 
to the gas phase due to the higher vapour pressure of the liquid phase compared to the crystalline phase. After the onset 
of a steady-state gas population, there are three growth channels for the crystal. Direct crystallization is the process by which 
crystals grow by incorporating nearby liquid particles. In the Bergeron process, liquid droplets first evaporate and the resulting gas phase 
contributes to the growth of crystalline regions. Ostwald ripening is instead the process by which small crystallites sublimate, and colloidal
particles are transferred to larger nuclei. In order to assess the relative importance of these three growth channels, we analyse the trajectory, and, for every time frame, assign each particle a state between gas, liquid and crystal. In order to minimize short-term fluctuations, the state of each colloidal particle is time averaged for $50\,\tau_B$. We then measure the fraction of trajectories with different transition histories.
Direct crystallization accounts for 54\% of particles trajectories in which
gas or liquid particles transition to the crystal state without liquid evaporation or crystal de-sublimation. The Bergeron process accounts for 35\% of particle trajectories in which liquid particles
transition to the gas state before crystallizing. Ostwald ripening accounts for 11\% of particle trajectories in which crystal particles transition to the gas state before returning
to the crystal state. The different crystallization channels are depicted in the diagram of Fig.~\ref{fig:transitions}c.}

\textcolor{red}{
In Supplementary Information we confirm these results with a detailed analysis of transition probabilities, including a state for surface particles that are in contact.
to crystalline particles.
While the direct freezing of the fluid represents the major contribution to the nuclei growth, the kinetic path via the gas phase (gas$\rightarrow$crystal), also plays a crucial role,
especially in determining the morphology of the porous crystal, as we discussed above. In this context the Bergeron process~\cite{glickman2000glossary,morrison2012resilience}
plays a considerably more important role than Ostwald ripening.}

@@

The process of formation of ``crystal gels'' may be generic to many other 
systems. The requirements are (i) the presence of gas-liquid phase separation below the melting point of a crystal, (ii) 
weak or little frustration against crystallization (in our case, the use of monodisperse colloids), 
(iii) dynamical slowing down in a supercooled liquid state, which is necessary to induce viscoelastic phase separation leading to the formation 
of a network structure of the minority liquid phase, and (iv) the degree of supercooling is low enough to avoid a high rate of densification of the liquid phase 
leading to vitrification.   
Many monoatomic and single-component molecular systems can satisfy all these conditions in a certain range of the temperature and pressure. 
This can be seen, for example, by looking at the phase diagram of a Lennard-Jones liquid \cite{lodge1997brownian}, which represents many molecular systems 
without specific directional interactions.

@@

\begin{figure*}[!t]
 \centering
 \includegraphics{phase_separation.pdf}
 \caption{{\bf Different routes of viscoelastic phase separation.} 
{\bf a,} Phase diagram. Empty black triangles are experimental point showing a fluid behaviour. The other symbols are the experimental points showing gel behaviour, consistently used in all panels. The dotted lines are the gas-liquid tie lines determined experimentally, with arrow heads indicating measured gas and liquid compositions (see Supplementary Information on the procedure to extract the compositions from confocal images).  Gray areas are theoretical equilibrium fluid and crystal regions. The solid and dashed lines are the metastable gas-liquid binodal and spinodal respectively.
{\bf b,} Radius of gyration ($R_g$) as a function of cluster size ($s$) for all state points at percolation. The
 gray line represents the fractal dimension of the random gelation universality class.
{\bf c,} Structure factors for the late stages of the gelation process. The gray and black lines represents respectively random gelation and compact fractal dimensions.
{\bf d,} Late stage of the evolution of local volume fraction around colloidal particles having 12 nearest neighbours.
 }
 \label{fig:phase_separation}
\end{figure*}

@@

\begin{figure*}[!t]
 \centering
 \includegraphics[width=12cm]{fig2}
 % crystal_size.eps: 0x0 pixel, 300dpi, 0.00x0.00 cm, bb=(atend)
\caption{\textcolor{red}{{\bf Percolated network structures.} {\bf a:} gel; {\bf b:} crystal-gel. The gel state is obtained from the state
point $\phi\approx 0.33$ and $c_p=1.07$ mg/g, while the crystal-gel is obtained at the same $\phi$ but
at lower polymer concentration, $c_p=0.38$ mg/g. On the left, the network is represented as a continuous field that has
been obtained by applying a gaussian filter to the position of the particles (of width equal to the inter-particle distance) and
plotting the surface which bounds the highest 20 \% values of the field. We can clearly see not only the percolated nature of both networks, 
but also the difference in the smoothness of the network surface between amorphous and crystal gels. 
On the right, slabs of 5 particle diameter thickness are plotted,
with particles coloured with the same colour scheme as in Fig.~\ref{fig:transitions}: red (crystal), green (gas particles), white (liquid particles). 
We note that very few liquid particles remain and the network is almost crystalline.}} 
\label{fig:network}
\end{figure*}

@@

\begin{figure*}[!t]
 \centering
 \includegraphics[width=14cm]{fig4}
 % snapshots.eps: 0x0 pixel, 300dpi, 0.00x0.00 cm, bb=0 -1 2256 735
 \caption{\textcolor{red}{{\bf Crystal gel formation.}  
{\bf a,} Reconstructions from confocal coordinates at $\phi\approx 0.33$ and $c_p=0.38$ mg/g. Depth of view is $5\sigma$. Particles are drawn to scale and coloured according to their phase. Red: crystal; white: liquid; green: gas. 
{\bf b,} Fraction of particles for each phase. Gray vertical lines indicate the times shown in panel a. 
{\bf c,} Each particle during its trajectory changes between three states: gas, liquid and crystal. Direct crystallization accounts for 54\% of particles trajectories in which
gas or liquid particles transition to the crystal state without liquid evaporation or crystal de-sublimation. The Bergeron process accounts for 35\% of particle trajectories in which liquid particles
transition to the gas state before crystallizing. Ostwald ripening accounts for 11\% of particle trajectories in which crystal particle transition to the gas state before returning
to the crystal state.}}
\label{fig:transitions}
\end{figure*}

>>>>

To gain additional insight, in Fig.~\ref{fig:phase_separation}d we plot the average local volume fraction as a function of the number of neighbours.
\textcolor{blue}{The volume fraction is obtained by computing the Voronoi diagram of each configuration,
which uniquely assigns a volume to each colloidal particle.}
For all state points, we can see that the average volume fraction of a colloidal
particle increases with its number of neighbours. Approaching close packing (12 neighbours), we clearly see two families of curves. The state point
at $\phi\approx 0.33$ and $c_p=0.38$ mg/g reaches close packing at a volume fraction of around $\sim 68\%$, which is in reasonable agreement with the
volume fraction of the stable crystal phase~\cite{ilett1995phase}. All other state points, instead, reach close packing at a volume fraction $\sim 64\%$, which is
indeed close to the volume fraction of the attractive glass state~\cite{pham2002multiple}. 
%(\textbf{the absolute values of these volume fraction are not totally convincing, probably because of a not optimal estimate of the size of the colloids. We should run the code of Mathieau with size detection}). 
This result clearly indicates that the state point at $\phi\approx 0.33$ and $c_p=0.38$ mg/g follows a different arrest mechanism, in which phase separation
is arrested by crystallization and not by glassiness. Although our estimation of the absolute volume fraction may involves systematic errors of a few \%, the trend should be robust~\cite{Poon2012}.

@@

\textcolor{blue}{Next we analyse in detail the crystallization trajectory, distinguishing between the different processes responsible for the
crystallization of the phase separated network.
Supplementary Movie 2 shows the confocal microscopy video for the crystallizing sample at $\phi\approx 0.33$ and $c_p=0.38$ mg/g,
where the process of crystal growth by direct attachment of gas particles can be observed directly. 
From the trajectory, we extract slabs of 10 particle diameter thickness for configurations at different times, shown in Fig.~\ref{fig:transitions}a. 
Particles are coloured according to their phases (gas, liquid, and crystal). For the classification criteria see Supplementary Information. The figure shows the first nucleation 
events inside the liquid network (left panel). At the same time liquid regions that have not crystallized start evaporating, and the gas phase contributes to the growth of nuclei (middle panel). Finally the different nuclei coalesce (right panel). \textcolor{red}{Note that the network pattern changes its surface smoothness (??) between $t=2055\,\tau_B$ (left panel) and $t=5577,\tau_B$ (middle panel) (see, e.g., Fig2a): 
not all liquid branches of the network crystallize due to evaporation (?). The crystal-gel structure instead persists, as seen by comparing $t=5577,\tau_B$ (middle panel) and $t=12098,\tau_B$ (right panel). 　
{\bf (Hajime: this part is not so clear for me.)}}}

\textcolor{blue}{
Next we show in Fig.~\ref{fig:transitions}b the fraction of particles in each different phase for the crystallization trajectory at
$\phi\approx 0.33$ and $c_p=0.38$ mg/g, after the liquid-gas phase separation has occurred.
The process of crystallization is characterized 
by a steep decrease in liquid particles, as they transform into small crystalline nuclei inside the liquid domains. This decrease
is then accompanied by an increase in the fraction of gas particles: as the first crystals start to reach the gas phase, liquid particles evaporate 
to the gas phase due to the higher vapour pressure of the liquid phase compared to the crystalline phase. After the onset 
of a steady-state gas population, there are three growth channels for the crystal. Direct crystallization is the process by which 
crystals grow by incorporating nearby liquid particles. In the Bergeron process, liquid droplets first evaporate and the resulting gas phase 
contributes to the growth of crystalline regions. Ostwald ripening is instead the process by which small crystallites sublimate, and colloidal
particles are transferred to larger nuclei. In order to assess the relative importance of these three growth channels, we analyse the trajectory, and, for every time frame, assign each particle a state between gas, liquid and crystal. In order to minimize short-term fluctuations, the state of each colloidal particle is time averaged for $50\,\tau_B$. We then measure the fraction of trajectories with different transition histories.
Direct crystallization accounts for 54\% of particles trajectories in which
gas or liquid particles transition to the crystal state without liquid evaporation or crystal de-sublimation. The Bergeron process accounts for 35\% of particle trajectories in which liquid particles
transition to the gas state before crystallizing. Ostwald ripening accounts for 11\% of particle trajectories in which crystal particles transition to the gas state before returning
to the crystal state. The different crystallization channels are depicted in the diagram of Fig.~\ref{fig:transitions}c.}

\textcolor{blue}{
In Supplementary Information we confirm these results with a detailed analysis of transition probabilities, including a state for surface particles that are in contact 
to crystalline particles.
While the direct freezing of the fluid represents the major contribution to the nuclei growth, the kinetic path via the gas phase (gas$\rightarrow$crystal), also plays a crucial role,
especially in determining the morphology of the porous crystal, as we discussed above. In this context the Bergeron process~\cite{glickman2000glossary,morrison2012resilience}
plays a considerably more important role than Ostwald ripening.}

@@

The process of formation of ``crystal gels'' may be generic to many other 
systems. The requirements are (i) the presence of gas-liquid phase separation below the melting point of a crystal, (ii) 
weak or little frustration against crystallization (in our case, the use of monodisperse colloids), 
(iii) dynamical slowing down in a supercooled liquid state, which is necessary to induce viscoelastic phase separation leading to the formation 
of a network structure of the minority liquid phase, and (iv) the degree of supercooling is low enough to avoid a high rate of densification of the liquid phase 
leading to vitrification.  
Many monoatomic and single-component molecular systems can satisfy all these conditions in a certain range of the temperature and pressure. 
This can be seen, for example, by looking at the phase diagram of a Lennard-Jones liquid \cite{lodge1997brownian}, which represents many molecular systems 
without specific directional interactions. \textcolor{blue}{For monoatomic systems such as noble metals, condition (iii) may not be satisfied easily. 
To access glassy slow dynamics in a liquid phase, we may need a deep quench at a high pressure. However, we note that even without strong dynamic asymmetry 
due to glassiness, bicontinuous phase separation can take place between 35-85 volume \% of the liquid phase in ordinary gas/liquid phase separation; 
and, thus, porous crystalline structures can be formed, although a thin network structure may be difficult to be formed.}

@@

\begin{figure*}[!t]
 \centering
 \includegraphics{phase_separation.pdf}
 \caption{{\bf Different routes of viscoelastic phase separation.} 
{\bf a,} Phase diagram. Empty black triangles are experimental point showing a fluid behaviour. The other symbols are the experimental points showing gel behaviour, consistently used in all panels. \textcolor{blue}{The dotted lines are the gas-liquid tie lines determined experimentally, with arrow heads indicating measured gas and liquid compositions (see Supplementary Information on the procedure to extract the compositions from confocal images).  Gray areas are theoretical equilibrium fluid and crystal regions. The solid and dashed lines are the metastable gas-liquid binodal and spinodal respectively.}
{\bf b,} Radius of gyration ($R_g$) as a function of cluster size ($s$) for all state points at percolation. The
 gray line represents the fractal dimension of the random gelation universality class.
{\bf c,} Structure factors for the late stages of the gelation process. The gray and black lines represents respectively random gelation and compact fractal dimensions.
\textcolor{blue}{{\bf d,} Late stage of the evolution of local volume fraction around colloidal particles having 12 nearest neighbours.}
 }
 \label{fig:phase_separation}
\end{figure*}

@@

\begin{figure*}[!t]
 \centering
 \includegraphics[width=12cm]{fig2}
 % crystal_size.eps: 0x0 pixel, 300dpi, 0.00x0.00 cm, bb=(atend)
\caption{\textcolor{blue}{{\bf Percolated network structures.} {\bf a:} gel; {\bf b:} crystal-gel. The gel state is obtained from the state
point $\phi\approx 0.33$ and $c_p=1.07$ mg/g, while the crystal-gel is obtained at the same $\phi$ but
at lower polymer concentration, $c_p=0.38$ mg/g. On the left, the network is represented as a continuous field that has
been obtained by applying a gaussian filter to the position of the particles (of width equal to the inter-particle distance) and
plotting the surface which bounds the highest 20 \% values of the field. We can clearly see not only the percolated nature of both networks, 
but also the difference in the smoothness of the network surface between amorphous and crystal gels. 
On the right, slabs of 5 particle diameter thickness are plotted,
with particles coloured with the same colour scheme as in Fig.~\ref{fig:transitions}: red (crystal), green (gas particles), white (liquid particles). 
We note that very few liquid particles remain and the network is almost crystalline. 
\textcolor{red}{\bf Hajime: we need to put length and time information for each panel.} 
}} 
\label{fig:network}
\end{figure*}

@@

\begin{figure*}[!t]
 \centering
 \includegraphics[width=14cm]{fig4}
 % snapshots.eps: 0x0 pixel, 300dpi, 0.00x0.00 cm, bb=0 -1 2256 735
 \caption{\textcolor{blue}{{\bf Crystal gel formation.}  
{\bf a,} Reconstructions from confocal coordinates at $\phi\approx 0.33$ and $c_p=0.38$ mg/g. Depth of view is $5\sigma$. Particles are drawn to scale and coloured according to their phase. Red: crystal; white: liquid; green: gas. \textcolor{red}{\bf Hajime: we need to put length information for images.}
{\bf b,} Fraction of particles for each phase. Gray vertical lines indicate the times shown in panel a. 
{\bf c,} Each particle during its trajectory changes between three states: gas, liquid and crystal. Direct crystallization accounts for 54\% of particles trajectories in which
gas or liquid particles transition to the crystal state without liquid evaporation or crystal de-sublimation. The Bergeron process accounts for 35\% of particle trajectories in which liquid particles
transition to the gas state before crystallizing. Ostwald ripening accounts for 11\% of particle trajectories in which crystal particle transition to the gas state before returning
to the crystal state.}}
\label{fig:transitions}
\end{figure*}
